---
title: "Chap 3"
author: "MITCH Holman"
date: "2/13/2018"
output: html_document
---
```{r label = "setup", include = FALSE, message = FALSE}
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE, comment = NA, fig.align = "center", warning = FALSE, message = FALSE)
```


### Chapter 3. Bar Graphs

```{r}
library(ggplot2)
library(gcookbook)
library(gridExtra)
library(plyr) 
```

### Section 3.1. Making a Basic Bar Graph

```{r}
ggplot(pg_mean, aes (x = group, y = weight))+
  geom_bar(stat = "identity")
ggplot(pg_mean, aes(x=group, y=weight)) +
    geom_bar(stat="identity", fill="lightblue", colour="black")


```

```{r, echo = FALSE}

c <-ggplot(BOD, aes(x=factor(Time), y=demand)) + geom_bar(stat="identity")

r <-ggplot(BOD, aes(x=Time, y=demand)) + geom_bar(stat="identity")


grid.arrange(c, r, ncol=2)
```






### Section 3.2. Grouping Bars Together

```{r}

ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(position="dodge", stat="identity")

ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(position="dodge", colour="black", stat="identity") +
    scale_fill_brewer(palette="Pastel1")
ce <- cabbage_exp[1:5, ]  


ggplot(ce, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(position="dodge", colour="black", stat="identity") +
    scale_fill_brewer(palette="Pastel1")

```

### Section 3.3. Making a Bar Graph of Counts

```{r}
ggplot(diamonds, aes(x=cut)) + geom_bar()

ggplot(diamonds, aes(x=carat)) + geom_bar()
```

### Section 3.4. Using Colors in a Bar Graph

```{r}
upc <- subset(uspopchange, rank(Change)>40)
upc

ggplot(upc, aes(x=Abb, y=Change, fill=Region)) + geom_bar(stat="identity")

ggplot(upc, aes(x=reorder(Abb, Change), y=Change, fill=Region)) +
    geom_bar(stat="identity", colour="black") +
    scale_fill_manual(values=c("#669933", "#FFCC66")) +
    xlab("State")
```

### Section 3.5. Coloring Negative and Positive Bars Differently

```{r}
csub <- subset(climate, Source=="Berkeley" & Year >= 1900)
csub$pos <- csub$Anomaly10y >= 0
t <-ggplot(csub, aes(x=Year, y=Anomaly10y, fill=pos)) +
  geom_bar(stat="identity", position="identity")


e <- ggplot(csub, aes(x=Year, y=Anomaly10y, fill=pos)) +
    geom_bar(stat="identity", position="identity", colour="black", size=0.25) +
    scale_fill_manual(values=c("#CCEEFF", "#FFDDDD"), guide=FALSE)
grid.arrange(t,e)
```

### Section 3.6. Adjusting Bar Width and Spacing

```{r}
tt <- ggplot(pg_mean, aes(x=group, y=weight)) + geom_bar(stat="identity")
rr <- ggplot(pg_mean, aes(x=group, y=weight)) + geom_bar(stat="identity", width=0.5)
qq <- ggplot(pg_mean, aes(x=group, y=weight)) + geom_bar(stat="identity", width=1)
grid.arrange(tt,rr,qq , ncol =3)

op <- ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(stat="identity", width=0.5, position="dodge")
og <- ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(stat="identity", width=0.5, position=position_dodge(0.7))

grid.arrange(op,og , ncol = 2)


ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(stat="identity", width=0.5, position="dodge")
ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(stat="identity", width=0.5, position=position_dodge(0.7))

```

### Section 3.7. Making a Stacked Bar Graph

```{r}

ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(stat="identity")

ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(stat="identity") +
    guides(fill=guide_legend(reverse=TRUE))
```

### Section 3.8. Making a Proportional Stacked Bar Graph

```{r}

ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar, order=desc(Cultivar))) +
    geom_bar(stat="identity")
ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(stat="identity", colour="black") +
    guides(fill=guide_legend(reverse=TRUE)) +
    scale_fill_brewer(palette="Pastel1")
```

### Section 3.9. Adding Labels to a Bar Graph

```{r}
ce <- ddply(cabbage_exp, "Date", transform,
            percent_weight = Weight / sum(Weight) * 100)

ggplot(ce, aes(x=Date, y=percent_weight, fill=Cultivar)) +
    geom_bar(stat="identity")

ggplot(ce, aes(x=Date, y=percent_weight, fill=Cultivar)) +
    geom_bar(stat="identity", colour="black") +
    guides(fill=guide_legend(reverse=TRUE)) +
    scale_fill_brewer(palette="Pastel1")
```
```{r}
the <- ggplot(cabbage_exp, aes(x=interaction(Date, Cultivar), y=Weight)) +
    geom_bar(stat="identity") +
    geom_text(aes(label=Weight), vjust=1.5, colour="white")


you <- ggplot(cabbage_exp, aes(x=interaction(Date, Cultivar), y=Weight)) +
    geom_bar(stat="identity") +
    geom_text(aes(label=Weight), vjust=-0.2)

grid.arrange(the, you, ncol= 2)
```
```{r}
ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(stat="identity", position="dodge") +
    geom_text(aes(label=Weight), vjust=1.5, colour="white",
              position=position_dodge(.9), size=3)
```

```{r}





ce <- ddply(ce, "Date", transform, label_y=cumsum(Weight)-0.5*Weight)
ggplot(ce, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(stat="identity") +
    geom_text(aes(y=label_y, label=Weight), vjust=1.5, colour="white")

ggplot(ce, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(stat="identity") +
    geom_text(aes(y=label_y, label=Weight), colour="white")
ggplot(ce, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(stat="identity", colour="black") +
    geom_text(aes(y=label_y, label=paste(format(Weight, nsmall=2), "kg")),
              size=4) +
    guides(fill=guide_legend(reverse=TRUE)) +
    scale_fill_brewer(palette="Pastel1")
```


### Section 3.10. Making a Cleveland Dot Plot

```{r, fig.height=5, fig.width=4}

library(gcookbook)
tophit <- tophitters2001[1:25, ] 

ggplot(tophit, aes(x=avg, y=name)) + geom_point()

ggplot(tophit, aes(x=avg, y=reorder(name, avg))) +
    geom_point(size=3) +                        # Use a larger dot
    theme_bw() +
    theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.y = element_line(colour="grey60", linetype="dashed"))

```


```{r}
ggplot(tophit, aes(x=reorder(name, avg), y=avg)) +
    geom_point(size=3) +                        # Use a larger dot
    theme_bw() +
    theme(axis.text.x = element_text(angle=60, hjust=1),
          panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.x = element_line(colour="grey60", linetype="dashed"))
```

```{r, fig.height=5, fig.width=5}
nameorder <- tophit$name[order(tophit$lg, tophit$avg)]
tophit$name <- factor(tophit$name, levels=nameorder)

ggplot(tophit, aes(x=avg, y=name)) +
    geom_segment(aes(yend=name), xend=0, colour="grey50") +
    geom_point(size=3, aes(colour=lg)) +
    scale_colour_brewer(palette="Set1", limits=c("NL","AL")) +
    theme_bw() +
    theme(panel.grid.major.y = element_blank(),   
          legend.position=c(1, 0.55),            
          legend.justification=c(1, 0.5))


ggplot(tophit, aes(x=avg, y=name)) +
    geom_segment(aes(yend=name), xend=0, colour="grey50") +
    geom_point(size=3, aes(colour=lg)) +
    scale_colour_brewer(palette="Set1", limits=c("NL","AL"), guide=FALSE) +
    theme_bw() +
    theme(panel.grid.major.y = element_blank()) +
    facet_grid(lg ~ ., scales="free_y", space="free_y")



```





<hr>
END!

