---
title: "Data With Tidyverse"
author: "MITCH"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)

ilo_data <-read.csv("ilo_data.csv")
ilo_hourly_compensation <- read.csv("ilo_hourly_compensation.csv")
ilo_working_hours <- read.csv("ilo_working_hours.csv")

```

## Join the two data sets together



```{r Join the two data sets together}



# Join both data frames
ilo_data <- ilo_hourly_compensation %>%
  inner_join(ilo_working_hours, by = c("country", "year"))

# Count the resulting rows
ilo_data  %>% 
    count()

# Examine ilo_data
ilo_data 
```

## Change variable types

```{r Change variable types}
# Turn year into a factor
ilo_data <- ilo_data %>%
  mutate(year = as.factor(as.numeric(year)))

# Turn country into a factor
ilo_data <- ilo_data %>%
  mutate(country = as.factor(country))
```


## Filter the data for plotting



```{r Filter the data for plotting}




# Examine the European countries vector
european_countries <- c("Finland"     ,   "France"      ,   "Italy"       ,   "Norway"   ,     
 "Spain"     ,     "Sweden"      ,   "Switzerland"   , "United Kingdom",
  "Belgium"     ,   "Ireland"  ,      "Luxembourg"   ,  "Portugal"  ,    
 "Netherlands" ,  "Germany"   ,   "Hungary"   ,   "Austria"  , "Czech Rep")

# Only retain European countries
ilo_data <- ilo_data %>%
  filter(country %in% european_countries)
ilo_data
```


## Some summary statistics


```{r Some summary statistics}
# Examine the structure of ilo_data
str(ilo_data)

# Group and summarize the data
ilo_data %>%
  group_by(year) %>%
  summarize(mean_hourly_compensation = mean(hourly_compensation),
            mean_working_hours = mean(working_hours))
```


## A basic scatter plot



```{r A basic scatter plot}
# Filter for 2006
plot_data <- ilo_data %>%
  filter(year == 2006)

# Create the scatter plot
plot_data

ggplot(plot_data) +  geom_point(aes(x = working_hours, y = hourly_compensation))
```


## Add labels to the plot


```{r Add labels to the plot}
# Create the plot
ggplot(plot_data) +
  geom_point(aes(x = working_hours, y = hourly_compensation)) +
  # Add labels
  labs(
    x = "Working hours per week",
    y = "Hourly compensation",
    title = "The more people work, the less compensation they seem to receive",
    subtitle = "Working hours and hourly compensation in European countries, 2006",
    caption = "Data source: ILO, 2017"
  )
```


## Apply a default theme




```{r Apply a default theme}
# Save your current plot into a variable: ilo_plot
ilo_plot <- ggplot(plot_data) +
  geom_point(aes(x = working_hours, y = hourly_compensation)) +
  labs(
    x = "Working hours per week",
    y = "Hourly compensation",
    title = "The more people work, the less compensation they seem to receive",
    subtitle = "Working hours and hourly compensation in European countries, 2006",
    caption = "Data source: ILO, 2017"
  )

# Try out theme_minimal
ilo_plot +
  theme_minimal()

# Try out any other possible theme function
ilo_plot +
  theme_light() 
```


## Change the appearance of titles



```{r Change the appearance of titles}
ilo_plot <- ilo_plot +
  theme_minimal() +
  # Customize the "minimal" theme with another custom "theme" call
  theme(
    text = element_text(family = "Bookman"),
    title = element_text(color = "gray25"),
    plot.subtitle = element_text(size = 12),
    plot.caption = element_text(color = "gray30")
  )
  
# Render the plot object
ilo_plot
```


## Alter background color and add margins



```{r Alter background color and add margins}
ilo_plot +
  # "theme" calls can be stacked upon each other, so this is already the third call of "theme"
  theme(
    plot.background = element_rect(fill = "gray95"),
    plot.margin = unit(c(5, 10, 5, 10), units = "mm")
  )

```


## Prepare the data set for the faceted plot


```{r Prepare the data set for the faceted plot }
# Filter ilo_data to retain the years 1996 and 2006
ilo_data <- ilo_data %>%
  filter(year == "1996" | year == "2006")
```

## Add facets to the plot





```{r Add facets to the plot}
# Again, you save the plot object into a variable so you can save typing later on
ilo_plot <- ggplot(ilo_data, aes(x = working_hours, y = hourly_compensation)) +
  geom_point() +
   labs(
    x = "Working hours per week",
    y = "Hourly compensation",
    title = "The more people work, the less compensation they seem to receive",
    subtitle = "Working hours and hourly compensation in European countries, 2006",
    caption = "Data source: ILO, 2017"
  ) +
  # Add facets here
  facet_grid(facets = . ~ year)
 
ilo_plot
```



## Define your own theme function



```{r Define your own theme function}
# For a starter, let's look at what you did before: adding various theme calls to your plot object
ilo_plot +
  theme_minimal() +
  theme(
    text = element_text(family = "Bookman", color = "gray25"),
    plot.subtitle = element_text(size = 12),
    plot.caption = element_text(color = "gray30"),
    plot.background = element_rect(fill = "gray95"),
    plot.margin = unit(c(5, 10, 5, 10), units = "mm")
  )
      
# Define your own theme function below
theme_ilo <- function() {
  theme_minimal() +
  theme(
    text = element_text(family = "Bookman", color = "gray25"),
    plot.subtitle = element_text(size = 12),
    plot.caption = element_text(color = "gray30"),
    plot.background = element_rect(fill = "gray95"),
    plot.margin = unit(c(5, 10, 5, 10), units = "mm")
  )
}
```


## Apply the new theme function to the plot


```{r Apply the new theme function to the plot}
# Apply your theme function
ilo_plot <- ilo_plot +
  theme_ilo()

# Examine ilo_plot
ilo_plot

ilo_plot +
   # Add another theme call
  theme(
    # Change the background fill to make it a bit darker
    strip.background = element_rect(fill = "gray60", color = "gray95"),
    # Make text a bit bigger and change its color
    strip.text = element_text(size = 11, color = "white")
  )
```


## A basic dot plot



```{r A basic dot plot}

# Create the dot plot
ggplot(ilo_data) +
  geom_path(aes(x = working_hours, y = country))

```


## Add arrows to the lines in the plot


```{r Add arrows to the lines in the plot}
ggplot(ilo_data) +
  geom_path(aes(x = working_hours, y = country),
  # Add an arrow to each path
            arrow = arrow(length = unit(1.5, "mm"), type = "closed"))
```


## Add some labels to each country


```{r Add some labels to each country}
ggplot(ilo_data) +
  geom_path(aes(x = working_hours, y = country),
            arrow = arrow(length = unit(1.5, "mm"), type = "closed")) +
  # Add a geom_text() geometry
  geom_text(
          aes(x = working_hours,
              y = country,
              label = round(working_hours, 1))
        )
```


## Reordering elements in the plot


```{r Reordering elements in the plot}
library(forcats)

# Reorder country factor levels
ilo_data <- ilo_data %>%
  # Arrange data frame
  arrange(year) %>%
  # Reorder countries by working hours in 2006
  mutate(country = fct_reorder(country,
                               working_hours,
                               last))

# Plot again
ggplot(ilo_data) +
  geom_path(aes(x = working_hours, y = country),
            arrow = arrow(length = unit(1.5, "mm"), type = "closed")) +
    geom_text(
          aes(x = working_hours,
              y = country,
              label = round(working_hours, 1))
          )
```


## Correct ugly label positions


```{r Correct ugly label positions}
# Save plot into an object for reuse
ilo_dot_plot <- ggplot(ilo_data) +
  geom_path(aes(x = working_hours, y = country),
            arrow = arrow(length = unit(1.5, "mm"), type = "closed")) +
    # Specify the hjust aesthetic with a conditional value
    geom_text(
          aes(x = working_hours,
              y = country,
              label = round(working_hours, 1),
              hjust = ifelse(year == "2006", 1.4, -0.4)
            ),
          # Change the appearance of the text
          size = 3,
          family = "Bookman",
          color = "gray25"
          )

ilo_dot_plot
```


## Change the viewport so labels don't overlap with plot border


```{r Change the viewport so }
# Reuse ilo_dot_plot
ilo_dot_plot <- ilo_dot_plot +
  # Add labels to the plot
  labs(
    x = "Working hours per week",
    y = "Country",
    title = "People work less in 2006 compared to 1996",
    subtitle = "Working hours in European countries, development since 1996",
    caption = "Data source: ILO, 2017"
  ) +
  # Apply your theme
  theme_ilo() +
  # Change the viewport
  coord_cartesian(xlim = c(25, 41))
  
# View the plot
ilo_dot_plot
```


## Optimizing the plot for mobile devices



```{r Optimizing the plot for mobile devices}
# Compute temporary data set for optimal label placement
median_working_hours <- ilo_data %>%
  group_by(country) %>%
  summarize(median_working_hours_per_country = median(working_hours)) %>%
  ungroup()

# Have a look at the structure of this data set
str(median_working_hours)

ilo_dot_plot +
  # Add label for country
  geom_text(data = median_working_hours,
            aes(y = country,
                x = median_working_hours_per_country,
                label = country),
            vjust = 2,
            family = "Bookman",
            color = "gray25") +
  # Remove axes and grids
  theme(
    axis.ticks = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    panel.grid = element_blank(),
    # Also, let's reduce the font size of the subtitle
    plot.subtitle = element_text(size = 9)
  )
```


## Introduction to the RMarkdown exercise interface

```{}
---
title: "Communicating with Data in the Tidyverse" 
author: "DataCamp"
date: "March 2018"
output: html_document
---
```

The following plot shows the relationship between the number of cylinders and miles per gallon in the `mtcars` dataset.

```{r}
boxplot(mtcars$mpg ~ mtcars$cyl)
```


As one can see, the more cylinders, the less miles per gallons. 








## Give your document a structure


```{}
---
title: "The reduction in weekly working hours in Europe" 
subtitle: "Looking at the development between 1996 and 2006"
author: "Insert your name here"
output: html_document
---
```
### Summary 

This is where you are going to quickly summarize what you did.

### Preparations 

This is where you will load the necessary R packages.

### Analysis

#### Data

This is where you will load the data. 

#### Preprocessing

This is where you will preprocess the data.

#### Results

This is where you will plot the data. 




## Change formatting of text snippets



```{}
---
title: "The reduction in weekly working hours in Europe" 
subtitle: "Looking at the development between 1996 and 2006"
author: "Mitch Holman"
output: html_document
---
```
### Summary 

The **International Labour Organization** (ILO) has many [data sets](http://www.ilo.org/global/statistics-and-databases/lang--en/index.htm) on working conditions. For example, one can look at how weekly working hours have been decreasing in many countries of the world, while monetary compensation has risen. In this report, *the reduction in weekly working hours* in European countries is analysed, and a comparison between 1996 and 2006 is made. All analysed countries have seen a decrease in weekly working hours since 1996 – some more than others.

### Preparations 

This is where you will load the necessary R packages.

###Analysis

#### Data

This is where you will load the data. 

#### Preprocessing

This is where you will preprocess the data.

###Results

This is where you will plot the data. 





## Specify packages in the first code chunk

```{}
---
title: "The reduction in weekly working hours in Europe" 
subtitle: "Looking at the development between 1996 and 2006"
author: "Insert your name here"
output: html_document
---
```

```{r}


```






## R code chunk options


```{}
---
title: "The reduction in weekly working hours in Europe" 
subtitle: "Looking at the development between 1996 and 2006"
author: "Insert your name here"
output: html_document
---
```






## Inline code statements

```{}
---
title: "The reduction in weekly working hours in Europe" 
subtitle: "Looking at the development between 1996 and 2006"
author: "Insert your name here"
output: html_document
---
```




## Add auxiliary images

```{}
---
title: "The reduction in weekly working hours in Europe" 
subtitle: "Looking at the development between 1996 and 2006"
author: "Insert your name here"
output: html_document
---
```






## Change the overall appearance of your report

```{}
---
title: "The reduction in weekly working hours in Europe" 
subtitle: "Looking at the development between 1996 and 2006"
author: "Insert your name here"
output: 
  html_document:
    theme: cosmo
    highlight: monochrome
---
```








## Add a table of contents


```{}
---
title: "The reduction in weekly working hours in Europe" 
subtitle: "Looking at the development between 1996 and 2006"
author: "Insert your name here"
output: 
  html_document:
    theme: cosmo
    highlight: monochrome
    toc: true
    toc_float: false
    toc_depth: 4
---
```







## More YAML hacks




```{}
---
title: "The reduction in weekly working hours in Europe" 
subtitle: "Looking at the development between 1996 and 2006"
author: "Insert your name here"
output: 
  html_document:
    theme: cosmo
    highlight: monochrome
    toc: true
    toc_float: false
    toc_depth: 4
    number_sections: true
    code_folding: hide
---
```






## Change style attributes of text elements


```{}
---
title: "The reduction in weekly working hours in Europe" 
subtitle: "Looking at the development between 1996 and 2006"
author: "Insert your name here"
output: 
  html_document:
    theme: cosmo
    highlight: monochrome
    toc: true
    toc_float: false
    toc_depth: 4
    code_folding: hide
---


```



## Reference the style sheet

```{}
---
title: "The reduction in weekly working hours in Europe" 
subtitle: "Looking at the development between 1996 and 2006"
author: "Insert your name here"
output: 
  html_document:
    theme: cosmo
    highlight: monochrome
    toc: true
    toc_float: false
    toc_depth: 4
    code_folding: hide
    css: styles.css
---
```






## Beautify a table with kable


```{}
---
title: "The reduction in weekly working hours in Europe" 
subtitle: "Looking at the development between 1996 and 2006"
author: "Insert your name here"
output: 
  html_document:
    theme: cosmo
    highlight: monochrome
    toc: true
    toc_float: false
    toc_depth: 4
    code_folding: hide
    css: styles.css
    df_print: kable
---
```


