---
title: "R Graphics Cookbook Chapter 9"
author: "Mitch Holman"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(ggplot2)
library(gcookbook)
```

## Controlling the Overall Appearance of Graphs

Note: Some problems cannot be completed as written because the code has changed. You will need to find the updated code.

### Section 9.1. Setting the Title of a Graph

```{r}
 p <- ggplot(heightweight, aes(x=ageYear, y=heightIn)) + geom_point()

p + ggtitle("Age and Height of Schoolchildren")

p + ggtitle("Age and Height\nof Schoolchildren")

p + ggtitle("Age and Height of Schoolchildren") +
    theme(plot.title=element_text(vjust = -2.5))

p + annotate("text", x=mean(range(heightweight$ageYear)), y=Inf,
             label="Age and Height of Schoolchildren", vjust=1.5, size=6)







```

### Section 9.2. Changing the Appearance of Text

```{r}
p <- ggplot(heightweight, aes(x=ageYear, y=heightIn)) + geom_point()


p + theme(axis.title.x=element_text(size=16, lineheight=.9, family="Times", face="bold.italic", colour="red"))

p + ggtitle("Age and Height\nof Schoolchildren") +
    theme(plot.title=element_text(size=rel(1.5), lineheight=.9, family="Times", face="bold.italic", colour="red"))



 p + annotate("text", x=15, y=53, label="Some text", size = 7, family="Times", fontface="bold.italic", colour="red")

p + geom_text(aes(label=weightLb), size=4, family="Times", colour="red")


```

### Section 9.3. Using Themes

```{r}
p <- ggplot(heightweight, aes(x=ageYear, y=heightIn)) + geom_point()


p + theme_grey()


p + theme_bw()

p + theme_grey(base_size=16, base_family="Times")


theme_set(theme_bw())

p

theme_set(theme_grey())




```

### Section 9.4. Changing the Appearance of Theme Elements

```{r}
p <- ggplot(heightweight, aes(x=ageYear, y=heightIn, colour=sex)) + geom_point()


p + theme(
    panel.grid.major = element_line(colour="red"),
    panel.grid.minor = element_line(colour="red", linetype="dashed", size=0.2),
    panel.background = element_rect(fill="lightblue"),
    panel.border = element_rect(colour="blue", fill=NA, size=2))


p + ggtitle("Plot title here") +
    theme(
    axis.title.x = element_text(colour="red", size=14),
    axis.text.x  = element_text(colour="blue"),
    axis.title.y = element_text(colour="red", size=14, angle = 90),
    axis.text.y  = element_text(colour="blue"),
    plot.title = element_text(colour="red", size=20, face="bold"))


p + theme(
    legend.background = element_rect(fill="grey85", colour="red", size=1),
    legend.title = element_text(colour="blue", face="bold", size=14),
    legend.text = element_text(colour="red"),
    legend.key = element_rect(colour="blue", size=0.25))


p + facet_grid(sex ~ .) + theme(
    strip.background = element_rect(fill="pink"),
    strip.text.y = element_text(size=14, angle=-90, face="bold"))
  
p + theme(axis.title.x = element_text(colour="red")) + theme_bw()


p + theme_bw() + theme(axis.title.x = element_text(colour="red", size=12))





```

### Section 9.5. Creating Your Own Themes

```{r}
mytheme <- theme_bw() +
    theme(text       = element_text(colour="red"),
          axis.title = element_text(size = rel(1.25)))

p <- ggplot(heightweight, aes(x=ageYear, y=heightIn)) + geom_point()

p + mytheme


```

### Section 9.6. Hiding Grid Lines

```{r}
p <- ggplot(heightweight, aes(x=ageYear, y=heightIn)) + geom_point()

p + theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank())


p + theme(panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank())


p + theme(panel.grid.major.y = element_blank(),
          panel.grid.minor.y = element_blank())



```

<hr>
END!