---
title: "R Graphics Cookbook Chapter 8"
author: "Mitch Holman"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(ggplot2)
library(gcookbook)
library(MASS) 
library(plyr) 
library(scales)
```

## Axes

Note: Some problems cannot be completed as written because the code has changed. You will need to find the updated code.

### Section 8.1. Swapping X- and Y- Axes

```{r}
 ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot()

ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() + coord_flip()


ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() + coord_flip() +
    scale_x_discrete(limits=rev(levels(PlantGrowth$group)))
```

### Section 8.2. Setting the Range of a Continuous Axis

```{r}
p <- ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot()

p + ylim(0, max(PlantGrowth$weight))

p + ylim(0, 10) + scale_y_continuous(breaks=NULL)

p + scale_y_continuous(breaks=NULL) + ylim(0, 10)

p + scale_y_continuous(limits=c(0, 10), breaks=NULL)

p + scale_y_continuous(limits = c(5, 6.5)) 

p + coord_cartesian(ylim = c(5, 6.5))

p + expand_limits(y=0)


```

### Section 8.3. Reversing a Continuous Axis

```{r}
ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() + scale_y_reverse()

ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() + ylim(6.5, 3.5)

ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() +
    scale_y_reverse(limits=c(8, 0))















```

### Section 8.4. Changing the Order of Items on a Categorial Axis

```{r}
p <- ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot()

p + scale_x_discrete(limits=c("trt1","ctrl","trt2"))

p + scale_x_discrete(limits=c("ctrl","trt1"))

p + scale_x_discrete(limits=rev(levels(PlantGrowth$group)))




```

### Section 8.5. Setting the the Scaling Ratio of the X- and Y-Axes

```{r}
sp <- ggplot(marathon, aes(x=Half,y=Full)) + geom_point()

sp + coord_fixed()

sp + coord_fixed() +
    scale_y_continuous(breaks=seq(0, 420, 30)) +
    scale_x_continuous(breaks=seq(0, 420, 30))

sp + coord_fixed(ratio=1/2) +
    scale_y_continuous(breaks=seq(0, 420, 30)) +
    scale_x_continuous(breaks=seq(0, 420, 15))









```

### Section 8.6. Setting the Position of Tick Marks

```{r}
ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot()

ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() +
    scale_y_continuous(breaks=c(4, 4.25, 4.5, 5, 6, 8))

ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() +
    scale_x_discrete(limits=c("trt2", "ctrl"), breaks="ctrl")







```

### Section 8.7. Removing Tick Marks and Labels

```{r}
p <- ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot()

p + theme(axis.text.y = element_blank())

p + theme(axis.ticks = element_blank(), axis.text.y = element_blank())

p + scale_y_continuous(breaks=NULL)











```

### Section 8.8. Changing the Text of Tick Labels

```{r}
hwp <- ggplot(heightweight, aes(x=ageYear, y=heightIn)) +
          geom_point()

hwp

hwp + scale_y_continuous(breaks=c(50, 56, 60, 66, 72),
                         labels=c("Tiny", "Really\nshort", "Short",
                                  "Medium", "Tallish"))

footinch_formatter <- function(x) {
    foot <- floor(x/12)
    inch <- x %% 12
    return(paste(foot, "'", inch, "\"", sep=""))
}

hwp + scale_y_continuous(labels=footinch_formatter)

hwp + scale_y_continuous(breaks=seq(48, 72, 4), labels=footinch_formatter)

timeHMS_formatter <- function(x) {
    h <- floor(x/60)
    m <- floor(x %% 60)
    s <- round(60*(x %% 1))                   
    lab <- sprintf("%02d:%02d:%02d", h, m, s)  
    lab <- gsub("^00:", "", lab)           
    lab <- gsub("^0", "", lab)            
    return(lab)
}







```

### Section 8.9. Changing the Appearance of Tick Labels

```{r}
bp <- ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() +
      scale_x_discrete(breaks=c("ctrl", "trt1", "trt2"),
                       labels=c("Control", "Treatment 1", "Treatment 2"))
bp

bp + theme(axis.text.x = element_text(angle=90, hjust=1, vjust=.5))

bp + theme(axis.text.x = element_text(angle=30, hjust=1, vjust=1))

bp + theme(axis.text.x = element_text(family="Times", face="italic",
                                      colour="darkred", size=rel(0.9)))













```

### Section 8.10. Changing the Text of Axis Labels

```{r}
hwp <- ggplot(heightweight, aes(x=ageYear, y=heightIn, colour=sex)) +
       geom_point()
hwp

hwp + xlab("Age in years") + ylab("Height in inches")

hwp + labs(x = "Age in years", y = "Height in inches")

hwp + scale_x_continuous(name="Age in years")

hwp + scale_x_continuous(name="Age\n(years)")










```

### Section 8.11. Removing Axis Labels

```{r}
p <- ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot()

p + theme(axis.title.x=element_blank())

p + xlab("")






```

### Section 8.12. Changing the Appearance of Axis Labels

```{r}
hwp <- ggplot(heightweight, aes(x=ageYear, y=heightIn)) + geom_point()

hwp + theme(axis.title.x=element_text(face="italic", colour="darkred", size=14))

hwp + ylab("Height\n(inches)") +
    theme(axis.title.y=element_text(angle=0, face="italic", size=14))

hwp + ylab("Height\n(inches)") +
    theme(axis.title.y=element_text(angle=90, face="italic", colour="darkred", size=14))









```

### Section 8.13. Showing Lines Along the Axis

```{r}
p <- ggplot(heightweight, aes(x=ageYear, y=heightIn)) + geom_point()

p + theme(axis.line = element_line(colour="black"))

p + theme_bw() +
    theme(panel.border = element_blank(),
          axis.line = element_line(colour="black"))

p + theme_bw() +
    theme(panel.border = element_blank(),
          axis.line = element_line(colour="black", size=4))

p + theme_bw() +
    theme(panel.border = element_blank(),
          axis.line = element_line(colour="black", size=4, lineend="square"))








```

### Section 8.14. Using a Logarhithmic Axis

```{r}


p <- ggplot(Animals, aes(x=body, y=brain, label=rownames(Animals))) +
     geom_text(size=3)
p

p + scale_x_log10() + scale_y_log10()

p + scale_x_log10(breaks=10^(-1:5)) + scale_y_log10(breaks=10^(0:3))

p + scale_x_log10(breaks=10^(-1:5),labels=trans_format("log10",math_format(10^.x))) +
    scale_y_log10(breaks=10^(0:3),labels=trans_format("log10", math_format(10^.x)))

ggplot(Animals, aes(x=log10(body), y=log10(brain), label=rownames(Animals))) +
    geom_text(size=3)

p + scale_x_continuous(trans = log_trans(),
                       breaks = trans_breaks("log", function(x) exp(x)),
                       labels = trans_format("log", math_format(e^.x))) +
    scale_y_continuous(trans = log2_trans(),
                       breaks = trans_breaks("log2", function(x) 2^x),
                       labels = trans_format("log2", math_format(2^.x)))

ggplot(aapl, aes(x=date,y=adj_price)) + geom_line()

ggplot(aapl, aes(x=date,y=adj_price)) + geom_line() +
    scale_y_log10(breaks=c(2,10,50,250))




```

### Section 8.15. Adding Ticks for a Logarhithmic Axis

```{r}
ggplot(Animals, aes(x=body, y=brain, label=rownames(Animals))) +
    geom_text(size=3) +
    annotation_logticks() +
    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
                  labels = trans_format("log10", math_format(10^.x))) +
    scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                  labels = trans_format("log10", math_format(10^.x)))


ggplot(Animals, aes(x=body, y=brain, label=rownames(Animals))) +
    geom_text(size=3) +
    annotation_logticks() +
    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
                  labels = trans_format("log10", math_format(10^.x)),
                  minor_breaks = log10(5) + -2:5) +
    scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                  labels = trans_format("log10", math_format(10^.x)),
                  minor_breaks = log10(5) + -1:3) +
    coord_fixed() +
    theme_bw()


```

### Section 8.16. Making a Circular Graph

```{r}
ggplot(wind, aes(x=DirCat, fill=SpeedCat)) +
    geom_histogram(binwidth=15, origin=-7.5) +
    coord_polar() +
    scale_x_continuous(limits=c(0,360))


ggplot(wind, aes(x=DirCat, fill=SpeedCat)) +
    geom_histogram(binwidth=15, origin=-7.5, colour="black", size=.25) +
    guides(fill=guide_legend(reverse=TRUE)) +
    coord_polar() +
    scale_x_continuous(limits=c(0,360), breaks=seq(0, 360, by=45),
                       minor_breaks=seq(0, 360, by=15)) +
    scale_fill_brewer()

coord_polar(start=-45 * pi / 180)

md <- data.frame(deaths = as.numeric(mdeaths),
                 month  = as.numeric(cycle(mdeaths)))


md <- ddply(md, "month", summarise, deaths = mean(deaths))
md

p <- ggplot(md, aes(x=month, y=deaths)) + geom_line() +
     scale_x_continuous(breaks=1:12)

p + coord_polar()

p + coord_polar() + ylim(0, max(md$deaths))

p + coord_polar() + ylim(0, max(md$deaths)) + xlim(0, 12)

mdx <- md[md$month==12, ]
mdx$month <- 0
mdnew <- rbind(mdx, md)

p %+% mdnew + coord_polar() + ylim(0, max(md$deaths))




```

### Section 8.17. Using Dates on an Axis

```{r}
ggplot(economics, aes(x=date, y=psavert)) + geom_line()


econ <- subset(economics, date >= as.Date("1992-05-01") &
                          date <  as.Date("1993-06-01"))


p <- ggplot(econ, aes(x=date, y=psavert)) + geom_line()
p


datebreaks <- seq(as.Date("1992-06-01"), as.Date("1993-06-01"), by="2 month")

p + scale_x_date(breaks=datebreaks) +
    theme(axis.text.x = element_text(angle=30, hjust=1))

p + scale_x_date(breaks=datebreaks, labels=date_format("%Y %b")) +
    theme(axis.text.x = element_text(angle=30, hjust=1))











```

### Section 8.18. Using Relative Times on an Axis

```{r}

www <- data.frame(minute = as.numeric(time(WWWusage)),
                  users  = as.numeric(WWWusage))


timeHM_formatter <- function(x) {
    h <- floor(x/60)
    m <- floor(x %% 60)
    lab <- sprintf("%d:%02d", h, m) 
    return(lab)
}


ggplot(www, aes(x=minute, y=users)) + geom_line()


ggplot(www, aes(x=minute, y=users)) + geom_line() +
    scale_x_continuous(name="time", breaks=seq(0, 100, by=10),
                       labels=timeHM_formatter)

scale_x_continuous(breaks=c(0, 20, 40, 60, 80, 100),
    labels=c("0:00", "0:20", "0:40", "1:00", "1:20", "1:40"))

timeHMS_formatter <- function(x) {
    h <- floor(x/3600)
    m <- floor((x/60) %% 60)
    s <- round(x %% 60)                    
    lab <- sprintf("%02d:%02d:%02d", h, m, s) 
    lab <- sub("^00:", "", lab)               
    lab <- sub("^0", "", lab)                 
    return(lab)
}










```

<hr>
END!