---
title: "R Graphics Cookbook Chapter 12"
author: "Mitch Holman"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(ggplot2)
library(gcookbook)
library(scales)
```

## Using Colors in Plots

Note: Some problems cannot be completed as written because the code has changed. You will need to find the updated code.

### Section 12.1. Setting the Colors in Objects

```{r}
 ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point(colour="red")

library(MASS) # For the data set
ggplot(birthwt, aes(x=bwt)) + geom_histogram(fill="red", colour="black")





```

### Section 12.2. Mapping Variables to Colors

```{r}
ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) 

ggplot(cabbage_exp, aes(x=Date, y=Weight)) 


ggplot(mtcars, aes(x=wt, y=mpg, colour=cyl)) + geom_point()

ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point(aes(colour=cyl))

ggplot(mtcars, aes(x=wt, y=mpg, colour=factor(cyl))) + geom_point()


m <- mtcars              
m$cyl <- factor(m$cyl)   
ggplot(m, aes(x=wt, y=mpg, colour=cyl)) + geom_point()

```

### Section 12.3. Using a Different Palette for a Discrete Variable

```{r}
p <- ggplot(uspopage, aes(x=Year, y=Thousands, fill=AgeGroup)) + geom_area()

p
p + scale_fill_discrete()
p + scale_fill_hue()

p + scale_fill_brewer()

h <- ggplot(heightweight, aes(x=ageYear, y=heightIn, colour=sex)) +
    geom_point()

h

h + scale_colour_hue(l=45)

library(RColorBrewer)
display.brewer.all()

p + scale_fill_brewer(palette="Oranges")

p + scale_fill_grey()

p + scale_fill_grey(start=0.7, end=0)





```

### Section 12.4. Using a Manually Defined Palette for a Discrete Variable

```{r}
h <- ggplot(heightweight, aes(x=ageYear, y=heightIn, colour=sex)) + geom_point()

h + scale_colour_manual(values=c("red", "blue"))

h + scale_colour_manual(values=c("#CC6666", "#7777DD"))

h + scale_colour_manual(values=c(m="blue", f="red"))






 
```

### Section 12.5. Using a Colorblind-Friendly Palette

```{r}
p <- ggplot(uspopage, aes(x=Year, y=Thousands, fill=AgeGroup)) + geom_area()

cb_palette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", 
                "#0072B2", "#D55E00", "#CC79A7")

p + scale_fill_manual(values=cb_palette)


c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00",
  "#CC79A7")


```

### Section 12.6. Using a Manually Defined Palette for a Continuous Variable

```{r}
p <- ggplot(heightweight, aes(x=ageYear, y=heightIn, colour=weightLb)) +
     geom_point(size=3)
p

p + scale_colour_gradient(low="black", high="white")


p + scale_colour_gradient2(low=muted("red"), mid="white", high=muted("blue"),
    midpoint=110)

p + scale_colour_gradientn(colours = c("darkred", "orange", "yellow", "white"))



```

### Section 12.7. Coloring a Shaded Region Based on Value

```{r}
cb <- subset(climate, Source=="Berkeley")

cb$valence[cb$Anomaly10y >= 0] <- "pos"
cb$valence[cb$Anomaly10y < 0]  <- "neg"

cb

interp <- approx(cb$Year, cb$Anomaly10y, n=1000)


cbi <- data.frame(Year=interp$x, Anomaly10y=interp$y)
cbi$valence[cbi$Anomaly10y >= 0] <- "pos"
cbi$valence[cbi$Anomaly10y < 0]  <- "neg"


ggplot(cbi, aes(x=Year, y=Anomaly10y)) +
    geom_area(aes(fill=valence), alpha = .4) +
    geom_line() +
    geom_hline(yintercept=0) +
    scale_fill_manual(values=c("#CCEEFF", "#FFDDDD"), guide=FALSE) +
    scale_x_continuous(expand=c(0, 0))








```

<hr>
END!