---
title: "R Graphics Cookbook Chapter 6"
author: "Mitch HOlman"
date: "`r Sys.Date()`"
output: html_document
---

### Chapter 6. Summarized Data Distributions

```{r}
library(ggplot2)
library(gcookbook)
library(MASS) 
library(plyr)

```

### Section 6.1. Making a Basic Histogram

```{r}
 
ggplot(faithful, aes(x=waiting)) + geom_histogram()

w <- faithful$waiting

ggplot(NULL, aes(x=w)) + geom_histogram()

ggplot(faithful, aes(x=waiting)) +
    geom_histogram(binwidth=5, fill="white", colour="black")


binsize <- diff(range(faithful$waiting))/15
ggplot(faithful, aes(x=waiting)) +
    geom_histogram(binwidth=binsize, fill="white", colour="black")

h <- ggplot(faithful, aes(x=waiting))  # Save the base object for reuse

h + geom_histogram(binwidth=8, fill="white", colour="black", origin=31)

h + geom_histogram(binwidth=8, fill="white", colour="black", origin=35)











```

### Section 6.2. Making Multiple Histograms from Grouped Data

```{r}

ggplot(birthwt, aes(x=bwt)) + geom_histogram(fill="white", colour="black") +
    facet_grid(smoke ~ .)


birthwt1 <- birthwt # Make a copy of the data

# Convert smoke to a factor
birthwt1$smoke <- factor(birthwt1$smoke)
levels(birthwt1$smoke)




library(plyr) # For the revalue() function
birthwt1$smoke <- revalue(birthwt1$smoke, c("0"="No Smoke", "1"="Smoke"))


 ggplot(birthwt1, aes(x=bwt)) + geom_histogram(fill="white", colour="black") +
    facet_grid(smoke ~ .)


 ggplot(birthwt, aes(x=bwt)) + geom_histogram(fill="white", colour="black") +
    facet_grid(race ~ .)


 ggplot(birthwt, aes(x=bwt)) + geom_histogram(fill="white", colour="black") +
    facet_grid(race ~ ., scales="free")



birthwt1$smoke <- factor(birthwt1$smoke)


ggplot(birthwt1, aes(x=bwt, fill=smoke)) +
    geom_histogram(position="identity", alpha=0.4)


```

### Section 6.3. Making a Density Curve

```{r}
ggplot(faithful, aes(x=waiting)) + geom_density()

ggplot(faithful, aes(x=waiting)) + geom_line(stat="density") +
    expand_limits(y=0)

w <- faithful$waiting

ggplot(NULL, aes(x=w)) + geom_density()

ggplot(faithful, aes(x=waiting)) +
    geom_line(stat="density", adjust=.25, colour="red") +
    geom_line(stat="density") +
    geom_line(stat="density", adjust=2, colour="blue")

ggplot(faithful, aes(x=waiting)) +
    geom_density(fill="blue", alpha=.2) +
    xlim(35, 105)


ggplot(faithful, aes(x=waiting)) +
    geom_density(fill="blue", colour=NA, alpha=.2) +
    geom_line(stat="density") +
    xlim(35, 105)

ggplot(faithful, aes(x=waiting, y=..density..)) +
    geom_histogram(fill="cornsilk", colour="grey60", size=.2) +
    geom_density() +
    xlim(35, 105)


```

### Section 6.4. Making Multiple Density Curves from Grouped Data

```{r}

birthwt1 <- birthwt


birthwt1$smoke <- factor(birthwt1$smoke)


ggplot(birthwt1, aes(x=bwt, colour=smoke)) + geom_density()


ggplot(birthwt1, aes(x=bwt, fill=smoke)) + geom_density(alpha=.3) 

ggplot(birthwt1, aes(x=bwt)) + geom_density() + facet_grid(smoke ~ .)


birthwt1$smoke <- revalue(birthwt1$smoke, c("0"="No Smoke", "1"="Smoke"))


 ggplot(birthwt1, aes(x=bwt)) + geom_density() + facet_grid(smoke ~ .)

ggplot(birthwt1, aes(x=bwt, y=..density..)) +
    geom_histogram(binwidth=200, fill="cornsilk", colour="grey60", size=.2) +
    geom_density() +
    facet_grid(smoke ~ .)








```

### Section 6.5. Making a Frequency Polygon

```{r}
ggplot(faithful, aes(x=waiting)) + geom_freqpoly()

ggplot(faithful, aes(x=waiting)) + geom_freqpoly(binwidth=4)



binsize <- diff(range(faithful$waiting))/15
ggplot(faithful, aes(x=waiting)) + geom_freqpoly(binwidth=binsize)







```

### Section 6.6. Making a Basic Box Plot

```{r}

ggplot(birthwt, aes(x=factor(race), y=bwt)) + geom_boxplot()

ggplot(birthwt, aes(x=factor(race), y=bwt)) + geom_boxplot(width=.5)

ggplot(birthwt, aes(x=factor(race), y=bwt)) +
    geom_boxplot(outlier.size=1.5, outlier.shape=21)


 ggplot(birthwt, aes(x=1, y=bwt)) + geom_boxplot() +
    scale_x_continuous(breaks=NULL) +
    theme(axis.title.x = element_blank())

 
 
 
 
 

```

### Section 6.7. Adding Notches to a Box Plot

```{r}

ggplot(birthwt, aes(x=factor(race), y=bwt)) + geom_boxplot(notch=TRUE)









```

### Section 6.8. Adding Means to a Box Plot

```{r}
ggplot(birthwt, aes(x=factor(race), y=bwt)) + geom_boxplot() +
    stat_summary(fun.y="mean", geom="point", shape=23, size=3, fill="white")
```

### Section 3.9. Making a Violin Plot

```{r}
p <- ggplot(heightweight, aes(x=sex, y=heightIn))

p + geom_violin()

p + geom_violin() + geom_boxplot(width=.1, fill="black", outlier.colour=NA) +
    stat_summary(fun.y=median, geom="point", fill="white", shape=21, size=2.5)

p + geom_violin(trim=FALSE)

p + geom_violin(scale="count")

p + geom_violin(adjust=2)

p + geom_violin(adjust=.5)



```

### Section 6.10. Making a Dot Plot

```{r}
countries2009 <- subset(countries, Year==2009 & healthexp>2000)

p <- ggplot(countries2009, aes(x=infmortality))

p + geom_dotplot()

p + geom_dotplot(binwidth=.25) + geom_rug() +
    scale_y_continuous(breaks=NULL) +   
    theme(axis.title.y=element_blank())

p + geom_dotplot(method="histodot", binwidth=.25) + geom_rug() +
    scale_y_continuous(breaks=NULL) + theme(axis.title.y=element_blank())

p + geom_dotplot(binwidth=.25, stackdir="center")
    scale_y_continuous(breaks=NULL) + theme(axis.title.y=element_blank())

p + geom_dotplot(binwidth=.25, stackdir="centerwhole")
    scale_y_continuous(breaks=NULL) + theme(axis.title.y=element_blank())




```

### Section 6.11. Making Multiple Dot Plots for Grouped Data

```{r}
ggplot(heightweight, aes(x=sex, y=heightIn)) +
    geom_dotplot(binaxis="y", binwidth=.5, stackdir="center")

ggplot(heightweight, aes(x=sex, y=heightIn)) +
    geom_boxplot(outlier.colour=NA, width=.4) +
    geom_dotplot(binaxis="y", binwidth=.5, stackdir="center", fill=NA)


ggplot(heightweight, aes(x=sex, y=heightIn)) +
    geom_boxplot(aes(x=as.numeric(sex) + .2, group=sex), width=.25) +
    geom_dotplot(aes(x=as.numeric(sex) - .2, group=sex), binaxis="y",
                 binwidth=.5, stackdir="center") +
    scale_x_continuous(breaks=1:nlevels(heightweight$sex),
                       labels=levels(heightweight$sex))






```

### Section 6.12. Making a Density Plot of Two-Dimensional Data

```{r}
p <- ggplot(faithful, aes(x=eruptions, y=waiting))

p + geom_point() + stat_density2d()

p + stat_density2d(aes(colour=..level..))

p + stat_density2d(aes(fill=..density..), geom="raster", contour=FALSE)


p + geom_point() +
    stat_density2d(aes(alpha=..density..), geom="tile", contour=FALSE)

p + stat_density2d(aes(fill=..density..), geom="raster", 
                   contour=FALSE, h=c(.5,5))








```


<hr>
END!