---
title: "R Graphics Cookbook Chapter 4"
author: "Mitchell Holman"
date: "`r Sys.Date()`"
output: html_document
---

```{r label = "setup", include = FALSE, message = FALSE}
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE, comment = NA, fig.align = "center", warning = FALSE, message = FALSE)
```


### Chapter 4. Line Graphs

```{r}
library(ggplot2)
library(gcookbook)
library(plyr)
library(gridExtra)
```

### Section 4.1. Making a Basic Line Plot

```{r}
ggplot(BOD, aes(x=Time, y=demand)) + geom_line()

BOD1 <- BOD 
BOD1$Time <- factor(BOD1$Time)
ggplot(BOD1, aes(x=Time, y=demand, group=1)) + geom_line()


ggplot(BOD, aes(x=Time, y=demand)) + geom_line() + ylim(0, max(BOD$demand))
ggplot(BOD, aes(x=Time, y=demand)) + geom_line() + expand_limits(y=0)
```

### Section 4.2. Adding Point to a Line Graph

```{r}
ggplot(BOD, aes(x=Time, y=demand)) + geom_line() + geom_point()


nok <- ggplot(worldpop, aes(x=Year, y=Population)) + geom_line() + geom_point()


vro <- ggplot(worldpop, aes(x=Year, y=Population)) + geom_line() + geom_point() +
    scale_y_log10()


grid.arrange(nok, vro)
```

### Section 4.3. Making a Line Graph with Multiple Lines

```{r}
tg <- ddply(ToothGrowth, c("supp", "dose"), summarise, length=mean(len))


sddd <-ggplot(tg, aes(x=dose, y=length, colour=supp)) + geom_line()


sdd <- ggplot(tg, aes(x=dose, y=length, linetype=supp)) + geom_line()

grid.arrange(sddd,sdd, ncol= 2)


ggplot(tg, aes(x=factor(dose), y=length, colour=supp, group=supp)) + geom_line()
```


```{r}
ggplot(tg, aes(x=dose, y=length)) + geom_line()
```

```{r}
qwe <- ggplot(tg, aes(x=dose, y=length, shape=supp)) + geom_line() +
    geom_point(size=4)         

wee <- ggplot(tg, aes(x=dose, y=length, fill=supp)) + geom_line() +
    geom_point(size=4, shape=21)
grid.arrange(qwe,wee, ncol= 2)

ggplot(tg, aes(x=dose, y=length, shape=supp)) +
    geom_line(position=position_dodge(0.2)) +  
    geom_point(position=position_dodge(0.2), size=4) 
```


### Section 4.4. Changing the Appearance of Lines

```{r}
ggplot(BOD, aes(x=Time, y=demand)) +
    geom_line(linetype="dashed", size=1, colour="blue")

tg <- ddply(ToothGrowth, c("supp", "dose"), summarise, length=mean(len))

ggplot(tg, aes(x=dose, y=length, colour=supp)) +
    geom_line() +
    scale_colour_brewer(palette="Set1")
```


```{r, fig.height=3, fig.width=7}
fog <- ggplot(tg, aes(x=dose, y=length, group=supp)) +
    geom_line(colour="darkgreen", size=1.5)


jik <- ggplot(tg, aes(x=dose, y=length, colour=supp)) +
    geom_line(linetype="dashed") +
    geom_point(shape=22, size=3, fill="white")

grid.arrange(fog, jik, ncol= 2)
```

### Section 4.5. Changing the Appearance of Points

```{r}
ggplot(BOD, aes(x=Time, y=demand)) +
    geom_line() +
    geom_point(size=4, shape=22, colour="darkred", fill="pink")

ggplot(BOD, aes(x=Time, y=demand)) +
    geom_line() +
    geom_point(size=4, shape=21, fill="white")

```

```{r}
tg <- ddply(ToothGrowth, c("supp", "dose"), summarise, length=mean(len))


pd <- position_dodge(0.2)

ggplot(tg, aes(x=dose, y=length, fill=supp)) +
    geom_line(position=pd) +
    geom_point(shape=21, size=3, position=pd) +
    scale_fill_manual(values=c("black","white"))
```


### Section 4.6. Making a Graph with Shaded Areas

```{r}
sunspotyear <- data.frame(
    Year     = as.numeric(time(sunspot.year)),
    Sunspots = as.numeric(sunspot.year)
)

ggplot(sunspotyear, aes(x=Year, y=Sunspots)) + geom_area()

sunspotyear <- data.frame(
    Year     = as.numeric(time(sunspot.year)),
    Sunspots = as.numeric(sunspot.year)
)

ggplot(sunspotyear, aes(x=Year, y=Sunspots)) + geom_area()

ggplot(sunspotyear, aes(x=Year, y=Sunspots)) +
    geom_area(colour="black", fill="blue", alpha=.2)

ggplot(sunspotyear, aes(x=Year, y=Sunspots)) +
    geom_area(fill="blue", alpha=.2) +
    geom_line()

```

### Section 4.7. Making a Stacked Area Graph

```{r}
ggplot(uspopage, aes(x=Year, y=Thousands, fill=AgeGroup)) + geom_area()

ggplot(uspopage, aes(x=Year, y=Thousands, fill=AgeGroup)) +
    geom_area(colour="black", size=.2, alpha=.4) +
    scale_fill_brewer(palette="Blues", breaks=rev(levels(uspopage$AgeGroup)))

ggplot(uspopage, aes(x=Year, y=Thousands, fill=AgeGroup, order=desc(AgeGroup))) +
    geom_area(colour="black", size=.2, alpha=.4) +
    scale_fill_brewer(palette="Blues")

ggplot(uspopage, aes(x=Year, y=Thousands, fill=AgeGroup, order=desc(AgeGroup))) +
    geom_area(colour=NA, alpha=.4) +
    scale_fill_brewer(palette="Blues") +
    geom_line(position="stack", size=.2)
```

### Section 4.8. Making a Proportional Stacked Area Graph

```{r}
uspopage_prop <- ddply(uspopage, "Year", transform,
                       Percent = Thousands / sum(Thousands) * 100)

ggplot(uspopage_prop, aes(x=Year, y=Percent, fill=AgeGroup)) +
    geom_area(colour="black", size=.2, alpha=.4) +
    scale_fill_brewer(palette="Blues", breaks=rev(levels(uspopage$AgeGroup)))

uspopage_prop <- ddply(uspopage, "Year", transform,
                       Percent = Thousands / sum(Thousands) * 100)

```

### Section 4.9. Adding a Confidence Region

```{r}
clim <- subset(climate, Source == "Berkeley",
               select=c("Year", "Anomaly10y", "Unc10y"))

ggplot(clim, aes(x=Year, y=Anomaly10y)) +
    geom_ribbon(aes(ymin=Anomaly10y-Unc10y, ymax=Anomaly10y+Unc10y),
                alpha=0.2) +
    geom_line()

ggplot(clim, aes(x=Year, y=Anomaly10y)) +
    geom_line(aes(y=Anomaly10y-Unc10y), colour="grey50", linetype="dotted") +
    geom_line(aes(y=Anomaly10y+Unc10y), colour="grey50", linetype="dotted") +
    geom_line()


```


<hr>
END!









